---
title: "Geog533 Lab 8"
author: "Xiaoxu Pan"
output:
  html_document:
    toc: yes
    toc_float: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}
RSS <- 2.3-1.7
n <- 24
df1 <- 1
df2 <- n-2
df <- n-1
RGMS <- RSS/df1
RSMS <- 1.7/df2
MT <- 2.3/df
F <- RGMS/RSMS
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|     `r RSS`     |      `r df1`       |    `r RGMS`    |      `r F`      |
| Residual |       1.7       |      `r df2`       |    `r RSMS`    |                 |
| Total    |       2.3       |      `r df`        |    `r MT`      |                


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
Premoney <- 0.46+0.19*50
Premoney
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?
Coefficient 0.19 represents a positive correlation between shopping trip frequency and annual income, which means that for every increase in annual income, the number of shopping trip for the individual will rise by 0.19.


### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
f.critical <- qf(0.95,1,22)
if(F>f.critical){
  print("The regression coefficient is significantly different from zero")
}else{
   print("The regression coefficient is not significantly different from zero")
}
```



### Question 1(e)
What is the value of the correlation coefficient?
```{r}
sr <- RSS/2.3
r <- sqrt(sr)
r
```





## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
snowfall <- c(36,78,11,45)
elevation <- c(400,800,200,675)
ms <- mean(snowfall)
me <- mean(elevation)
ds <- snowfall-ms
de <- elevation-me
r <- sum(ds*de)/(3*sd(snowfall)*sd(elevation))
slope <- r*sd(snowfall)/sd(elevation)
intercept <- ms-slope*me
print(paste("The intercept is", round(intercept,digits = 5)))
print(paste("The slope is", round(slope,digits = 5)))
```


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
TSS <- sum(ds*ds)
RGSS <- TSS*r*r
RSSS <- TSS-RGSS
RSMS <- RSSS/2
rerror <- sqrt(RSMS)
rerror
```


### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
F <- RGSS/RSMS
f.critical <- qf(0.95,1,2)
if(F>f.critical){
  print("The regression coefficient associated with the independent variables is not equal to zero")
}else{
  print("The regression coefficient associated with the independent variables is equal to zero")
}
result <- t.test(snowfall,elevation)
print(paste("The 95% confidence interval is",result$conf.int[1],result$conf.int[2]))
```


### Question 2(d)
Find the value of $r^2$.
```{r}
r2 <- r*r
r2
```


### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
x <- c(400,800,200,675)
y <- intercept+slope*x
e <- snowfall-y
DF <- as.data.frame(cbind(x,y,e))
names(DF) <- c("observed values","predicted values","residuals")
DF
```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
df1 <- 1
df2 <- 4-2
df <- 4-1
F <- RGSS/RSMS
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|    `r RGSS`     |     `r df1`        |   `r RGSS`     |     `r F`       |
| Residual |    `r RSSS`     |     `r df2`        |   `r RSMS`     |                 |
| Total    |    `r TSS`      |     `r df`         |                |                 |

### Question 2(g)
Graph the data and the regression line.
```{r}
plot(snowfall~elevation)
m <- lm(snowfall~elevation)
abline(m,col="red",lwd=2)
```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
df <- read.csv(url)
y <- df$SalePrice
x <- df$LotSize
m <- lm(y~x)
summary(m)
print(paste("The regression correlation is y =",m$coefficients[1],"+",m$coefficients[2],"* x"))
```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
df <- read.csv(url)
y <- df$price
x <- df$bedrooms
m <- lm(y~x)
summary(m)
print(paste("The regression correlation is y =",m$coefficients[1],"+",m$coefficients[2],"* x"))
```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
x <- df$bathrooms
m <- lm(y~x)
summary(m)
print(paste("The regression correlation is y =",m$coefficients[1],"+",m$coefficients[2],"* x"))
print("There are positive correlation between the number of bathrooms and house price. The correlation could represent 1.4% of the relationship while the correlation of the number of bedrooms and house price is stronger than the correlation of the number of bathrooms and house price ")
```


## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
url <- "http://spatial.binghamton.edu/geog533/data/hometown.csv"
df <- read.csv(url)
newRow <- data.frame(lastname="Pan",firstname="Xiaoxu",type="grads",location="40.223034,116.237537",latitude="40.223034",longitude="116.237537",image="",fullname="Xiaoxu Pan",popup="")
dfnew <- rbind(df,newRow)
dfnew[30:34,]
if(!require(leaflet)) {install.packages("leaflet")}
library(leaflet)
m <- leaflet(width = "100%") %>%
  addTiles() %>%
  addMarkers(lng=116.237537, lat=40.223034,popup="Xiaoxu Pan")
m
```

