---
title: "Geog533 Lab 10"
author: "Xiaoxu Pan"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r}
if(!require(UScensus2010)) {install.packages("UScensus2010")}
library(UScensus2010)
if(!require(UScensus2010county)) {install.county("osx")}
library(UScensus2010county)
data("new_york.county10")
shp <- new_york.county10
plot(shp)
```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r}
if(!require(tmap)) {install.packages("tmap")}
library(tmap)
qtm(shp)
```


### Question 1(c)	
How many counties in New York State?
```{r}
df <- shp@data
nrow(df)
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
df$county[df$NAMELSAD10=="Broome County"]
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
if(!require(moments)) {install.packages("moments")}
library(moments)
population <- df$P0010001
sum(population)
s <- summary(population)
s[1]
s[3]
s[4]
s[6]
skewness(population)
```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(population)
boxplot(population)

```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)


### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
if(!require(UScensus2010tract)) {install.tract("osx")}
library(UScensus2010tract)
data("new_york.tract10")
shp <-new_york.tract10
plot(shp)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
df <- shp@data
sum(df$P0010001)
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
shp1 <- shp[new_york.tract10$county=="007",]
plot(shp1)
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
sum(shp1$P0010001)
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
pop <- shp1$P0010001
hist(pop)
boxplot(pop)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r}
library(rgdal)
df1 <- shp1@data
dfnew <- df1[,1:5]
dfnew$ratio <- dfnew$P0010001/sum(dfnew$P0010001)
dfnew
shp1@data <- dfnew
writeOGR(shp1,dsn = "output",layer = "shp1",driver="ESRI Shapefile",overwrite_layer = TRUE)
shpnew <- readOGR(dsn = "output", layer = "shp1")

```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)


### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
if(!require(ncdf4)) {install.packages("ncdf4")}
library(ncdf4)
library(raster)
ndvi <- brick("NDVI.nc")
ndvi

```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
nrow(ndvi)
ncol(ndvi)
ncell(ndvi)
nbands(ndvi)
res(ndvi)
extent(ndvi)
bbox(ndvi)
projection(ndvi)
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
Meanndvi <- mean(ndvi)
Maxndvi <- max(ndvi)

writeRaster(Meanndvi,filename = "Meanndvi.tif")
writeRaster(Maxndvi,filename = "Maxndvi.tif")

```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}
plot(ndvi,13:24)
```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r}
hist(ndvi,13:24)

```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r}
plot(ndvi,7)
#values <- click(ndvi, n=1, xy=FALSE)
values<-ndvi[50,60]
plot(as.vector(values), type="b" )

```

