---
title: "Lab4"
author: "Xiaoxu Pan"
date: "2017/10/6"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(raster)
require(rgdal)
```

## Question 1
The number of points scored by each team in a tournament is normally distributed, with mean = 32 and standard deviation = 7. What is the probability of

### Question1a 
A team scoring 20 points or fewer?
```{r}
pnorm(20,mean = 32,sd = 7)
curve(dnorm(x,32,7),xlim = c(0,64),main='Normal Distribution')
cord.x <- c(0,seq(0,20,by = 0.01),20)
cord.y <- c(0,dnorm(seq(0,20,0.01),mean = 32,sd = 7),0)
polygon(cord.x,cord.y,density = 50,col = 'red')
```

### Question1b
A team scoring more than 35 points?
```{r}
pnorm(35,mean = 32,sd = 7,lower.tail = FALSE)
curve(dnorm(x,32,7),xlim = c(0,64),main='Normal Distribution')
cord.x <- c(35,seq(35,64,by = 0.01),64)
cord.y <- c(0,dnorm(seq(35,64,0.01),mean = 32,sd = 7),0)
polygon(cord.x,cord.y,density = 50,col = 'red')
```

### Question1c
A team scoring between 20 and 40 points?
```{r}
diff(pnorm(c(20,40),mean = 32,sd = 7))
curve(dnorm(x,32,7),xlim = c(0,64),main='Normal Distribution')
cord.x <- c(20,seq(20,40,by = 0.01),40)
cord.y <- c(0,dnorm(seq(20,40,0.01),mean = 32,sd = 7),0)
polygon(cord.x,cord.y,density = 50,col = 'red')
```

## Question 2
The number of comments per post on a social media site is exponentially distributed, with the average post receiving ten comments. What percentage of posts get:

### Question2a
Fewer than three comments? 
```{r}
pexp(3,rate = 1/10)
curve(dexp(x,rate = 1/10),xlim = c(0,50),main='Exponential Distribution')
cord.x <- c(0,seq(0,3,0.01),3)
cord.y <- c(0,dexp(seq(0,3,0.01),rate = 1/10),0)
polygon(cord.x,cord.y,density = 50,col = 'red')
```

### Question2b
More than 20 comments?
```{r}
pexp(20,rate = 1/10,lower.tail = FALSE)
curve(dexp(x,rate = 1/10),xlim = c(0,50),main='Exponential Distribution')
cord.x <- c(20,seq(20,50,0.01),50)
cord.y <- c(0,dexp(seq(20,50,0.01),rate = 1/10),0)
polygon(cord.x,cord.y,density = 50,col = 'red')
```

### Question2c
Between five and ten comments?
```{r}
diff(pexp(c(5,10),rate = 1/10))
curve(dexp(x,rate = 1/10),xlim = c(0,50),main='Exponential Distribution')
cord.x <- c(5,seq(5,10,0.01),10)
cord.y <- c(0,dexp(seq(5,10,0.01),rate = 1/10),0)
polygon(cord.x,cord.y,density = 50,col = 'red')
```

## Question 3
Basic raster creation and calculation:

### Question3a
Create a first raster (30 rows 30 columns) and assign values to the raster based on random values from a uniform distribution.
```{r}
ras1 <- raster(nrows=30,ncols=30,xmn=0,xmx=30,ymn=0,ymx=30)
ras1[] <- runif(ncell(ras1))
plot(ras1)
hist(ras1)
```

### Question3b
Create a second raster (30 rows 30 columns) and assign values to the raster based on random values from a normal distribution.
```{r}
ras2 <- raster(nrows=30,ncols=30,xmn=0,xmx=30,ymn=0,ymx=30)
ras2[] <- rnorm(ncell(ras2))
plot(ras2)
hist(ras2)
```

### Question3c
Create a third raster (30 rows 30 columns) and assign values to the raster based on cell-by-cell addition of the first raster and the second raster. 
```{r}
ras3 <- raster(nrows=30,ncols=30,xmn=0,xmx=30,ymn=0,ymx=30)
ras3[] <- ras1[] + ras2[]
plot(ras3)
hist(ras3)
```

### Question3d
Calculate the mean value of the third raster. Reclassify the third raster into a binary image: 1 (cell value > mean value) and 0 (cell value <= mean value). Save the reclassified image to your working directory as a TIFF image.
```{r}
M <- cellStats(ras3,"mean")
M
ras3[ras3 > M] <- 1
ras3[ras3 <= M] <- 0
writeRaster(ras3,filename = "reclassifiedR.tif",overwrite = TRUE)
plot(ras3)
hist(ras3)
```

